Pool Model

### 2025-08-28 pool_model.tex
more outlook comment
format introduction

\section{Introduction}
Motivation: general, flexible approach for bacterial systems that allows easily modify the system depending on processes on different scales.
Allows us to describe the processes within populations using different bacterial states as well as interactions between different populations using ODEs.\\
Existing models in Predictive Microbiology (Primary Models)~\cite{perez-rodriguez_predictive_2012}:
\begin{itemize}
    \item An overview of primary models used in food science can be found in~\cite{van_boekel_kinetic_2008}.
    \item Baranyi-Roberts Model~\cite{baranyi_modeling_1993, baranyi_dynamic_1994}
    \item modified Gompertz
    \item Buchanan et al.: three-phase linear model for lag phase, exponential growth phase, and stationary phase~\cite{buchanan_when_1997}
    \item McKellar two-compartment model with bacteria in growing and non-growing state~\cite{mckellar_heterogeneous_1997}
    \item Inactivation models
\end{itemize}
Secondary models:
\begin{itemize}
    \item Polynomial models 
    \item Square Root-Type Models, e.g., Ratkowski for temperature~\cite{ratkowsky_relationship_1982}
    \item The Cardinal Parameter Model~\cite{zwietering_decision_1992}
\end{itemize}
Comprehensive reviews about competition/cooperation bacterial mechanisms: \cite{ghoul_ecology_2016, stubbendieck_bacterial_2016, hibbing_bacterial_2010, west_social_2007}\\

@@

% ------------------------------------------------------------------------------------------------
\section{Conclusion}
\noindent
\begin{itemize}
    \item New theoretical framework
    \item consistent with \ac{ode} models (Baranyi-Roberts)
    \item Better explainability
    \item Include effects such as inhibition, toxins in understandable manner
    \item showed spatial limitations of \ac{ode} model
    \item \ac{abm} model allowed modeling of cellular processes; ensure that pool model is derivable
        from the individual behvaiour of cellular processes
\end{itemize}

>>>>

\newpage
%
%
%
%
% ================================================================================================
\section{Introduction}
Motivation: general, flexible approach for bacterial systems that allows easily modify the system depending on processes on different scales.
Allows us to describe the processes within populations using different bacterial states as well as interactions between different populations using ODEs.

\paragraph{Bacteria Consortia and processes (growth, lag phase, inhibition, etc.)}
\begin{itemize}
    \item 
\end{itemize}

\noindent
\paragraph{ODE models (predictive microbiology, existing, overview, most popular)}
Existing models in Predictive Microbiology (Primary Models)~\cite{perez-rodriguez_predictive_2012}:
\begin{itemize}
    \item An overview of primary models used in food science can be found in~\cite{van_boekel_kinetic_2008}.
    \item Baranyi-Roberts Model~\cite{baranyi_modeling_1993, baranyi_dynamic_1994}
    \item modified Gompertz
    \item Buchanan et al.: three-phase linear model for lag phase, exponential growth phase, and stationary phase~\cite{buchanan_when_1997}
    \item McKellar two-compartment model with bacteria in growing and non-growing state~\cite{mckellar_heterogeneous_1997}
    \item Inactivation models
\end{itemize}
%
Secondary models:
\begin{itemize}
    \item Polynomial models 
    \item Square Root-Type Models, e.g., Ratkowski for temperature~\cite{ratkowsky_relationship_1982}
    \item The Cardinal Parameter Model~\cite{zwietering_decision_1992}
\end{itemize}
%
%
Comprehensive reviews about competition/cooperation bacterial mechanisms: \cite{ghoul_ecology_2016, stubbendieck_bacterial_2016, hibbing_bacterial_2010, west_social_2007}\\

\noindent
\paragraph{Spatial effects in cosortia (adhesion, migration, chemotaxis)}
\begin{itemize}
    \item 
\end{itemize}

\noindent
\paragraph{Models for spatial effects; PDEs, \acp{abm}, others}

@@

% ------------------------------------------------------------------------------------------------
\section{Conclusion}
\noindent
\begin{itemize}
    \item New theoretical framework
    \item consistent with \ac{ode} models (Baranyi-Roberts)
    \item Better explainability
    \item Include effects such as inhibition, toxins in understandable manner
    \item showed spatial limitations of \ac{ode} model
    \item \ac{abm} model allowed modeling of cellular processes; ensure that pool model is derivable
        from the individual behvaiour of cellular processes
    \item outlook: generalized mathematical formulation; extend inhibition model (because it is a
        dimensional reduction of existing models)
\end{itemize}
